---
title: "Practical Machine Learning Project"
author: "Derek Luo"
date: "2017.3.2"
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```
#**Study Background**
####Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

#**Executive Summary**
####We clean the data set and fit three models. The first model was built by Decision Tree, but the accuracy was only 52.06%. So we built the second model by Random Forest with highest accuracy  99.64%. The third model was built by Gradient Boosting Machine Model, which accuracy was 98.66%. We then select compare the last two models for the testing data sets, and get identical predictions. 




#**Data Processing**
```{r , message=F, warning=FALSE}
##Load the packages we need.
library(data.table)
library(caret)
library(ggplot2)

##Download the data if they don't exist in working directory.
if(!file.exists("./training.csv")){
url.training <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
download.file(url.training, destfile = "./training.csv")
}
if(!file.exists("./testing.csv")){
url.testing <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(url.testing, destfile = "./testing.csv")
}

##Read the data into our workspace.
training <- fread("./training.csv", na.strings=c("NA","", "#DIV/0!"))
testing <- fread("./testing.csv", na.strings=c("NA","", "#DIV/0!"))

```
* **We replace "" and "#DIV/0!" as NA when we notice they are missing values, too.**
* **There are 19,622 observations and 160 variables in the training data set**
* **There are 20 observations and 160 variables in the testing data set.**



#**Data Cleaning**

```{r , results = "hide", warning=FALSE}
training <- data.table(training)
testing <- data.table(testing)

##We drop the columns that only contains zero.
nacols <- colnames(testing)[colSums(!is.na(testing)) == 0]
training[,(nacols):= NULL]
testing[,(nacols):= NULL]

##We drop the index ,username and timestamp columns that we won't use in our analysis.
unwantedcols <- c("V1", "user_name", "raw_timestamp_part_1",
                  "raw_timestamp_part_2","cvtd_timestamp", "problem_id")
training[,(unwantedcols):= NULL]
testing[,(unwantedcols):= NULL]

##We set the variables into correct classes.
training$new_window <- as.factor(training$new_window)
testing$new_window <- as.factor(testing$new_window)
training$classe <- as.factor(training$classe)

#Set the seed for reproducible computation.
set.seed(12345)

##We split training data set into smaller training data set and validation data set.
intrain <- createDataPartition(training$classe, p = 0.7)
trainingdp <- training[c(unlist(intrain))]
validationdp <- training[-c(unlist(intrain))]
```



#**Model Building**
###**Decision Tree**
```{r}
##We set 10 folds for cross validation
train_control <- trainControl(method="cv", number=10)
##We fit our first model with decision tree.
modelFit1 <- train(classe ~., method="rpart", data=trainingdp,
trControl = train_control)
result1 <- confusionMatrix(validationdp$classe, predict(modelFit1, newdata=validationdp))
result1
```
#####**Somehow the accuracy is only 52.06%, which is not acceptable. So we just stop here and try other models.**



###**Random Forest**
```{r}
modelFit2 <- train(classe ~., method="rf", data=trainingdp)
m2final <- modelFit2$finalModel
m2final
result2 <- confusionMatrix(validationdp$classe, predict(modelFit2, newdata=validationdp))
result2

```

###**Gradient Boosting Machine**
```{r , message=F, warning=FALSE}
fitControl <- trainControl(method = "repeatedcv",
                           number = 5,
                           repeats = 1)

modelFit3 <- train(classe ~ ., data=trainingdp, method = "gbm",
                 trControl = fitControl,
                 verbose = FALSE)

result3 <- confusionMatrix(validationdp$classe, predict(modelFit3, newdata=validationdp))
result3



```


```{r}
ggplot(modelFit3)

```






###**Predict the "classe" for testing data.**
```{r}
##Prediction of Random Forest
testing2 <- predict(modelFit2, newdata = testing)
testing2

##Prediction of Generalized Boosted Model
testing3 <- predict(modelFit3, newdata = testing)
testing3

identical(testing2, testing3)
table(testing2, testing3)
##We can see the predictions are identical for both models.
```




